Sex-Based Differences in Tuberculosis Immune Response: The Role of Leptin

### 2025-03-15 products/manuscript/Manuscript.qmd
Discussion

>>>>

This study examined the role of leptin in metabolism and immune response regulation. The sample included 30 males and 30 females with the first TB episode from Uganda. Our findings revealed that females had higher body fat storage, higher leptin levels, and stronger immune responses. Body fat was positively associated with leptin levels, and leptin mediated the effect of body fat on immune activation.

Our findings align with previous studies showing that females tend to have a stronger immune response than males [@Klein2016]. This difference in immunity may influence susceptibility and the speed of infection clearance between sexes [@Takahashi2021]. This study focused on immune activation and identified a sex-based difference in CD4+ cell levels, consistent with findings from a previous study [@Bothamley2021]. Recent research also indicates that changes in body fat in males can significantly reduce CD8+ cell levels compared to females [@Dhanyalayam2024]. Our study, along with previous research, provides evidence of sex-specific differences in the initiation and dynamic response of the immune system to TB infection. Our study also identified leptin as a bridge between metabolism and the immune system. This suggests its potential as a therapeutic target for regulating immune responses based on specific metabolic conditions [@Naylor2016]. Animal-based preclinical studies have demonstrated its practical significance [@Fantuzzi2000], but further research is needed to assess its clinical applicability in human populations.

This study has several limitations. First, the small sample size reduced the statistical power of the analysis. Sample loss due to the instability of the deuterium dilution approach in measuring body water further limited the detection of significant associations. Second, non-biological factors such as professional exposure, alcohol use, and smoking status [@Wigger2022] were not controlled, which could have influenced immune responses. Given their higher prevalence among males [@Soh2017], these factors may have confounded the findings. Despite these limitations, this study provides valuable insights that complement previous research. Future studies should explore the dynamic regulatory role of leptin at different stages of immune responses.
